Selective stock enhancement undermines community stability

### 2022-01-24 rmd/manuscript.Rmd
Update manuscript.Rmd

# theory infomation
df_theory <- read_csv(here::here("result/result_ricker.csv"))

n_sim <- df_theory %>% 
  mutate(n_sim = n_warmup + n_burnin + n_timestep) %>% 
  pull(n_sim) %>% 
  unique()

param_set <- df_theory %>% 
  filter(r_max == 2,
         alpha != 0.3) %>% 
  distinct(n_species,
           k,
           r_type,
           r1,
           r_min,
           r_max,
           sd_env,
           phi,
           int_type,
           alpha,
           model,
           seed)

@@

We used a multispecies Ricker model [@fowlerSpeciesDynamicsAlter2012] to simulate community dynamics with selective stock enhancement of a constituent species (species 1). Specifically, population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

In our simulation, we used `r nrow(param_set)` simulation scenarios with varied values of the following ecological parameters: intrinsic growth of the enhanced species $r_1$, mean strength of interspecific competition $\bar{\alpha}$, carrying capacity $K$, relative fitness of captive-bred individuals $f_R$, and degree of environmental stochasticity $\sigma_{\epsilon}$. Under each scenario with a unique parameter combination, we ran `r n_sim` time steps of `r max(df_theory$n_rep)` independent communities (i.e., simulation replicates) with varying numbers of released individuals $R$ ($0-500$). The last `r unique(df_theory$n_timestep)` time steps were used to obtain the coefficient of variation, temporal mean, and temporal standard deviation of total community density ($\sum_i^SN_i$). Temporal means and standard deviations were calculated separately for the enhanced species $N_1$ and unenhanced species $\sum_{i,i \ne 1}^S N_i$ to infer mechanisms that underlie the community-level response.

In general, our model predicted a destabilizing effect of selective stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of released individuals $R$ (Figure XX). This pattern stemmed mainly from the reduced mean of total community density, and both enhanced and unenhanced species groups were responsible for this result. For enhanced species, released individuals induced overcompensation, in which the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals. Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals, confirming previous theoretical predictions [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition caused the reduced mean density of unenhanced species and led to decreased numbers of persisting species at high levels of stock enhancement. Combined, the long-term average of the total community density decreased more sharply than individual species groups with stock enhancement (XX). The SDs showed a similar trend, but the relationship was flatter at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density resulted in the increased CV. These patterns were qualitatively similar across most ecological scenarios, even when the interspecific competition is very weak (XX).

The destabilizing effect emerges because stock enhancement undermines the balance of species interactions that underpin community stability. Ecological theory suggests that stable coexistence of competing species requires a niche difference that is large enough to overcome the relative difference in intrinsic competitive ability [@carrollNicheFitnessDifferences2011a]. Under this condition, competing species can grow from small populations because numerically dominant species undergoes stronger intraspecific competition [@carrollNicheFitnessDifferences2011a]. Such stable coexistence increases the likelihood of stabilizing temporal dynamics of species rich communities [@tilmanEcologicalConsequencesChanges1999a] because it gives rise to "overyielding" [@loreauDoesFunctionalRedundancy2004; @carrollNicheFitnessDifferences2011a], that is, total community density of a multi-species community exceeds what would be expected in a single species community ($\sum_i^S N_i > K$). However, stock enhancement is externally controlled and the number of release is not subject to density-dependent regulation. Therefore, excessive individuals supplemented through stock enhancement impose additional intra- and interspecific competition that interferes with the ecological process producing overyielding.

The only scenario where stock enhancement had little influences on community dynamics was when carrying capacity was sufficiently large ($K=500$). In particular, stock enhancement increased, albeit slightly, the average density of the enhanced species with low population growth rate ($r_1=0.5$). This result is reasonable because some of the best evidence for successful stock enhancement comes from critically endangered species [@johnsonGeneticRestorationFlorida2010; @oppelPopulationReinforcementDemographic2021]. In the meantime, our theoretical analysis also illuminates limitation of stock enhancement as a primary tool of ecosystem management. Habitat loss and fragmentation have severely reduced biodiversity, and one of the primary motivations for stock enhancement is to mitigate or reverse such a trend. However, stock enhancement will not bring desired ecological outcomes unless we address the root causes of habitat degradation and restore environmental capacity to support biodiversity.

@@

Under each parameter combination, we ran `r n_sim` time steps of `r max(df_theory$n_rep)` independent communities (i.e., simulation replicates). The number of released individuals $R$ was drawn randomly from a uniform distribution for each simulation replicate as $R \sim Unif(0,500)$. We initialized the community with populations of each species drawn from a poisson distribution with a mean of five. We repeated the seeding procedure every 10 time steps over the first `r unique(df_theory$n_warmup)` time steps to allow species to establish populations and reach an equilibrium without stock enhancement (initialization period). After the initialization period, we released $R$ individuals of the enhanced species every time step over the next `r unique(df_theory$n_burnin)` time steps to reach a new equilibrium with selective stock enhancement (burn-in period). After the burn-in period, we continued the simulation run with stock enhancement and saved the last `r unique(df_theory$n_timestep)` time steps to obtain the relative temporal fluctuations (i.e., the coefficient of variation) of the whole community ($\sum_i^SN_j$), the enhanced species ($N_1$), and the unenhanced species ($\sum_{i,i \ne 1}^S N_i$). We also calculated temporal means and standard deviations for each species group. Values of simulation parameters were summarized in XXX. We performed simulations using R version XXX.

>>>>

# theory infomation
load(here::here("result/result_ricker.RData"))

n_sim <- sim_result %>% 
  mutate(n_sim = n_warmup + n_burnin + n_timestep) %>% 
  pull(n_sim) %>% 
  unique()

param_set <- sim_result %>% 
  filter(r_max == 2,
         alpha != 0.3) %>% 
  distinct(n_species,
           k,
           r_type,
           r1,
           r_min,
           r_max,
           sd_env,
           phi,
           int_type,
           alpha,
           model,
           seed)

@@

We employed a multispecies Ricker model [@fowlerSpeciesDynamicsAlter2012] to simulate community dynamics with selective stock enhancement of a constituent species (species 1). Specifically, population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

In our simulation, we used `r nrow(param_set)` simulation scenarios with varied values of the following ecological parameters: intrinsic growth of the enhanced species $r_1$, mean strength of interspecific competition $\bar{\alpha}$, carrying capacity $K$, relative fitness of captive-bred individuals $f_R$, and degree of environmental stochasticity $\sigma_{\epsilon}$. Under each scenario with a unique parameter combination, we ran `r n_sim` time steps of `r max(sim_result$n_rep)` independent communities (i.e., simulation replicates) with varying numbers of released individuals $R$ ($0-500$). The last `r unique(sim_result$n_timestep)` time steps were used to obtain the number of species persist (defined as $N_i > 0.01$ at $t=1000$), the coefficient of variation (CV), the temporal mean, and the temporal SD of total community density ($\sum_i^SN_i$). In addition, temporal means and SDs were calculated separately for the enhanced species $N_1$ and unenhanced species $\sum_{i,i \ne 1}^S N_i$ to infer mechanisms that underlie the community-level response.

In general, our model predicted a destabilizing effect of selective stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of released individuals $R$ (**Figure 1A**). This pattern stemmed mainly from the reduced mean of total community density, and both enhanced and unenhanced species groups were responsible for this result (**Figure 1A**). For enhanced species, released individuals induced overcompensation, in which the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals. Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals, confirming previous theoretical predictions [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition caused the reduced mean density of unenhanced species and led to decreased numbers of persisting species at high levels of stock enhancement (**Figure 1A**). Combined, the long-term average of the total community density decreased more sharply than individual species groups with stock enhancement (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density resulted in the increased CV. These patterns were qualitatively similar across most ecological scenarios, even when interspecific competition is very weak (**Figure S1-6**).

The destabilizing effect emerges because stock enhancement undermines the balance of species interactions that underpins community stability. Ecological theory suggests that stable coexistence of competing species requires a niche difference that is large enough to overcome the relative difference in intrinsic competitive ability [@carrollNicheFitnessDifferences2011a]. Under this condition, competing species can grow from small populations because numerically dominant species undergoes stronger intraspecific competition [@carrollNicheFitnessDifferences2011a]. Such stable coexistence increases the likelihood of stabilizing temporal dynamics of species rich communities [@tilmanEcologicalConsequencesChanges1999a] because it gives rise to "overyielding" [@loreauDoesFunctionalRedundancy2004; @carrollNicheFitnessDifferences2011a], that is, total community density of a multi-species community exceeds what would be expected in a single species community ($\sum_i^S N_i > K$). However, stock enhancement is externally controlled and the number of release is not subject to density-dependent regulation. Therefore, excessive individuals supplemented through stock enhancement impose additional intra- and interspecific competition that interferes with the ecological process producing overyielding.

The only scenario where stock enhancement had a little influence on community dynamics was when carrying capacity was sufficiently large ($K=500$; **Figures S7-S12**). In particular, stock enhancement increased, albeit slightly, the average density of the enhanced species with low population growth rate ($r_1=0.5$; **Figures S7 and S8**). This result may explain why some of the best evidence for successful stock enhancement comes from endangered species [@johnsonGeneticRestorationFlorida2010; @oppelPopulationReinforcementDemographic2021], which typically suffer from increased mortality. At the same time, our theoretical analysis illuminates limitation of stock enhancement as a primary tool of ecosystem management. One of the primary motivations for stock enhancement is to mitigate or reverse declining trends of biodiversity due to human impacts, such as habitat loss and fragmentation [@chaseEcosystemDecayExacerbates2020]. However, our results suggest that stock enhancement will not bring desired ecological outcomes unless we restore high levels of environmental capacity to support biodiversity.

@@

Under each parameter combination, we ran `r n_sim` time steps of `r max(sim_result$n_rep)` independent communities (i.e., simulation replicates). The number of released individuals $R$ was drawn randomly from a uniform distribution for each simulation replicate as $R \sim Unif(0,500)$. We initialized the community with populations of each species drawn from a poisson distribution with a mean of five. We repeated the seeding procedure every 10 time steps over the first `r unique(sim_result$n_warmup)` time steps to allow species to establish populations and reach an equilibrium without stock enhancement (initialization period). After the initialization period, we released $R$ individuals of the enhanced species every time step over the next `r unique(sim_result$n_burnin)` time steps to reach a new equilibrium with selective stock enhancement (burn-in period). After the burn-in period, we continued the simulation run with stock enhancement and saved the last `r unique(sim_result$n_timestep)` time steps to obtain the relative temporal fluctuations (i.e., the coefficient of variation) of the whole community ($\sum_i^SN_j$), the enhanced species ($N_1$), and the unenhanced species ($\sum_{i,i \ne 1}^S N_i$). We also calculated temporal means and standard deviations for each species group. Values of simulation parameters were summarized in XXX. We performed simulations using R version XXX.
